Classical probability measures as pullback of quantum probability measures?

### 2022-05-10 pullback.tex
pullback.lyx Global replace the subindex of the probability spaces along axes...

Since the previous abstraction doesn't specify the process to generate
the outcomes, this process could well be a quantum experiment. Let
us prepare a beam of one kind of spin~$1$ particles, and split it
by an inhomogeneous magnetic field along the $z$-axis as a well-known
Stern-Gerlach type experiment \cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum,schwinger2001quantum}.
This beam will be split into three beams, and we will call the outcome
beams bent up as $\ket{0}$, bent down as $\ket{1}$, and not-bent
as $\ket{2}$. Instead of sending a beam of particles, if only one
particle is sent to the beam splitter, we will only get one of the
outcomes $\ket{0}$, $\ket{1}$, and $\ket{2}$, and there is a probability
to get each outcome. In the language of our abstraction, the set of
outcomes is $\Omega_{0}=\{\ket{0},\ket{1},\ket{2}\}$. All possible
events are $\emptyset$, $\left\{ \ket{0}\right\} $, $\left\{ \ket{1}\right\} $,
$\left\{ \ket{2}\right\} $, $\left\{ \ket{0},\ket{1}\right\} $,
$\left\{ \ket{0},\ket{2}\right\} $, $\left\{ \ket{1},\ket{2}\right\} $,
and $\Omega_{0}$. Each time we perform this experiment, it induces
a classical probability measure $\mu_{0}\colon2^{\Omega_{0}}\rightarrow\left[0,1\right]$.
Moreover, if we prepare the particles in the ``same'' way as the
particle inducing $\mu_{0}$, they will induce the same classical
probability measure $\mu_{0}$.

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
sending the ``same'' particle as the previous one, but we rotate
the inhomogeneous magnetic field $90\lyxmathsym{\textdegree}$ and
now along the $x$-axis. This time we call the outcome beams bent
left as $\ket{\ps}$, bent right as $\ket{\ms}$, and not-bent still
as $\ket{2}$. Although the sample space $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $
and the probability measure $\mu_{1}\colon2^{\Omega_{1}}\rightarrow\left[0,1\right]$
defined by this experiment is different from the previous one, they
all produce a not-bent outcome~$\ket{2}$, and the probability of
the common event $\left\{ \ket{2}\right\} $ is believed to be the
same, i.e., 
\begin{equation}
\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)\,.\label{eq:mu0-equal-mu1}
\end{equation}

@@

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but represent
the same situation. For example, if we take the complement on both
sides of Eq.~(\ref{eq:mu0-equal-mu1}), we will have
\begin{equation}
\mu_{0}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{0}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{1}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This identification can be achieved by mapping a classical
event~$E$ to the projector generated by $E$,\footnote{Should we prove whether there is no other way to identify?}
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$, because
$\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)$ is equal to
$\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)$ as operators,
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}

>>>>

Since the previous abstraction doesn't specify the process to generate
the outcomes, this process could well be a quantum experiment. Let
us prepare a beam of one kind of spin~$1$ particles, and split it
by an inhomogeneous magnetic field along the $z$-axis as a well-known
Stern-Gerlach type experiment \cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum,schwinger2001quantum}.
This beam will be split into three beams, and we will call the outcome
beams bent up as $\ket{0}$, bent down as $\ket{1}$, and not-bent
as $\ket{2}$. Instead of sending a beam of particles, if only one
particle is sent to the beam splitter, we will only get one of the
outcomes $\ket{0}$, $\ket{1}$, and $\ket{2}$, and there is a probability
to get each outcome. In the language of our abstraction, the set of
outcomes is $\Omega_{z}=\{\ket{0},\ket{1},\ket{2}\}$. All possible
events are $\emptyset$, $\left\{ \ket{0}\right\} $, $\left\{ \ket{1}\right\} $,
$\left\{ \ket{2}\right\} $, $\left\{ \ket{0},\ket{1}\right\} $,
$\left\{ \ket{0},\ket{2}\right\} $, $\left\{ \ket{1},\ket{2}\right\} $,
and $\Omega_{z}$. Each time we perform this experiment, it induces
a classical probability measure $\mu_{z}\colon2^{\Omega_{z}}\rightarrow\left[0,1\right]$.
Moreover, if we prepare the particles in the ``same'' way as the
particle inducing $\mu_{z}$, they will induce the same classical
probability measure $\mu_{z}$.

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
sending the ``same'' particle as the previous one, but we rotate
the inhomogeneous magnetic field $90\lyxmathsym{\textdegree}$ and
now along the $x$-axis. This time we call the outcome beams bent
left as $\ket{\ps}$, bent right as $\ket{\ms}$, and not-bent still
as $\ket{2}$. Although the sample space $\Omega_{x}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $
and the probability measure $\mu_{x}\colon2^{\Omega_{x}}\rightarrow\left[0,1\right]$
defined by this experiment is different from the previous one, they
all produce a not-bent outcome~$\ket{2}$, and the probability of
the common event $\left\{ \ket{2}\right\} $ is believed to be the
same, i.e., 
\begin{equation}
\mu_{z}\left(\left\{ \ket{2}\right\} \right)=\mu_{x}\left(\left\{ \ket{2}\right\} \right)\,.\label{eq:mu0-equal-mu1}
\end{equation}

@@

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but represent
the same situation. For example, if we take the complement on both
sides of Eq.~(\ref{eq:mu0-equal-mu1}), we will have
\begin{equation}
\mu_{z}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{z}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{x}\left(\left\{ \ket{2}\right\} \right)=\mu_{x}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This identification can be achieved by mapping a classical
event~$E$ to the projector generated by $E$,\footnote{Should we prove whether there is no other way to identify?}
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$, because
$\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)$ is equal to
$\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)$ as operators,
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}
